---
title: "Hands-on exercise06"
author: "Yuheng Liang"
format: html
editor: visual
date: "September 16, 2024"
date-modified: "September 16,2024"
execute: 
  eval: true
  echo: true
  freeze: true
---

# Hands_on exercise06 Measures of Spatial Autocorrelation
## 1.0 Tools and data
### 1.1 setting the tools
spdep: will be used to compute spatial weights, global and local spatial autocorrelation statistics.
```{R}
pacman::p_load(sf,spdep,tmap,tidyverse,ggplot2)
```
### 1.2  Importing the data
importing the shapefile into environment
```{R}
hunan <- st_read(dsn = "data/geospatial", layer = "hunan")
```
importing csv file
```{R}
#| eval: false
hunan2012 <- read_csv("data/aspatial/hunan_2012.csv")
```
performing relational join
```{r}
#| eval: false
hunan <-left_join(hunan,hunan2012)%>%
  select(1:4,7,15)
```

```{R}
#| eval: false
write_rds(hunan,"data/rds/hunan.rds")
```

```{R}
hunan<- read_rds("data/rds/hunan.rds")
```
### 1.3 Visualising Region Development Indicator
preapre a basemap and a choropleth map showing the distribution of GDPPC 2012 by using qtm() of tmaps package.
#### 1.3.1 Equal interval classification
```{R}
equal <- tm_shape(hunan)+
  tm_fill("GDPPC",
          n=5,
          style ="equal" )+
  tm_borders(alpha = 0.5)+
  tm_layout(main.title = "Equal interval classification")
```

#### 1.3.2 Equal quantitle classification
```{R}
quantile <- tm_shape(hunan)+
  tm_fill("GDPPC",
          n=5,
          style = "quantile")+
  tm_borders(alpha = 0.5)+
  tm_layout(main.title = "Equal quantitle classification")
```

```{R}
tmap_arrange(equal,
             quantile,
             asp=1,
             ncol = 2)
```

## 2.0 Global Measures of Spatial Autocorrelation
### 2.1 computing contiguity spatial weights
Before we can compute the global spatial autocorrelation statistics, we need to construct a spatial weights of the area.
poly2nb() of spdep packages is used to compute contiguity weigh for the study area.
If you do not specify this argument the default is set to TRUE, that is, if you don’t specify queen = FALSE this function will return a list of first order neighbours using the Queen criteria.
```{R}
wm_q <- poly2nb(hunan, queen = TRUE)
summary(wm_q)
```
the report show that there are 88 area units in hunan. The most connected area unit has 11 neightsbours. There are two units with only one neighbours.
### 2.2 Row-standardised weights matrix
assign weights to each neighboring polygon.
```{r}
rswm_q <- nb2listw(wm_q, style = "W",zero.policy = TRUE)
rswm_q
```
* The input of nb2listw() must be an object of class nb. The syntax of the function has two major arguments,namely style and zero.poly
* style can be W,B,C,U,minimax,S
* If zero policy is set to true, weights vectors of zero length are inserted for regions without neighbour in the neighbours list.

## 3.0 Global Measures of Spatial Autocorrelation: Moran's 1
perform Moran's 1 statistics testing by using moran.test() of spdep
### 3.1 MAron's 1 test
Moran's statistical testing using moran.test() of spdep
```{R}
moran.test(hunan$GDPPC, 
           listw=rswm_q, 
           zero.policy = TRUE, 
           na.action=na.omit)
```
alternative hypothesis greater: This indicates that the alternative hypothesis is that the moran's 1 is greater than 0, meaning that postive spatial autocorrelation is expected

The result means that areas with high GDPPC tend to be clustered toghther, and areas with low GDPPC also tend to cluster together

### 3.2 Computing Monte Carlo Moran's 1
The code chunk below performs permutation test for Moran's 1 statistic by using moran.mc() of spdep. A total of 1000 simulation will be performed.
```{R}
set.seed(1234)
bperm= moran.mc(hunan$GDPPC,
                 listw = rswm_q,
                 nsim = 999,
                 zero.policy = TRUE,
                 na.action = na.omit)
bperm
```
The caculated Moran's value is 0.30075.The observed Moran's value has the highest rank among the 1000 simulated values, indicating it is the largest.The p-value is 0.001, which is very small.

The high Moran's value and the very low p-value indicate a significant positive spatial autocorrelation. This means that areas with high GDPPC tend to be clustered together, and areas with low GDPPC also tend to cluster together.

### 3.3 Visualising Monte Carlo Moran's I
plotting the distribution of the statistical values as a histogram by using the code chunk below.
```{R}
mean(bperm$res[1:999])
var(bperm$res[1:999])
summary(bperm$res[1:999])
```
In the code chunk below hist() and abline() of R Graphics are used.
```{R}
hist(bperm$res,
     freq = TRUE,
     breaks = 20,
     xlab = "Simulate Moran's I")
abline(v=0,col="red")
```
The simulated Moran's value are centered around a particular value with a relatively symmetrical spread and the majority of simulated Moran's valued are centered around 0 which is expected under the null hypothesis of no spatial autocorrelation


