---
title: "In-class_Ex10"
author: "Yuheng Liang"
format: html
editor: visual
date: "Oct 28, 2024"
date-modified: "Oct 28, 2024"
execute: 
  eval: true
  echo: true
  freeze: true
---

# In-class_exe 10

## Getting started

```{r}
pacman::p_load(olsrr,ggstatsplot,sf,tmap,tidyverse,gtsummary,performance,see,sfdep)
```

## importing the data

```{r}
condo_resale <- read_csv("data/aspatial/Condo_resale_2015.csv")
```

```{r}
mpsz <- read_rds("data/mpsz.rds")
```

```{r}
condo_resale_sf <- read_rds("data/condo_resale_sf.rds")
```



Building a Hedonic Pricing Model by using Multiple Linear Regression Method

```{r}
condo_mlr <- lm(formula = SELLING_PRICE ~ AREA_SQM + AGE    + 
                  PROX_CBD + PROX_CHILDCARE + PROX_ELDERLYCARE +
                  PROX_URA_GROWTH_AREA + PROX_HAWKER_MARKET + PROX_KINDERGARTEN + 
                  PROX_MRT  + PROX_PARK + PROX_PRIMARY_SCH + 
                  PROX_TOP_PRIMARY_SCH + PROX_SHOPPING_MALL + PROX_SUPERMARKET + 
                  PROX_BUS_STOP + NO_Of_UNITS + FAMILY_FRIENDLY + FREEHOLD, 
                data=condo_resale_sf)
```


Generating tidy linear regression report
```{r}
ols_regress(condo_mlr)
```

```{r}
olsrr_condo <- ols_regress(condo_mlr)
```

Variable selection
```{r}
condo_fw_mlr <- ols_step_forward_p(condo_mlr,
                                   p_val = 0.05,
                                   datails = FALSE)
condo_fw_mlr
```
plot
```{r}
plot(condo_fw_mlr)
```

```{r}
condo_bw_mlr <- ols_step_backward_p(condo_mlr,
                                   p_val = 0.05,
                                   datails = FALSE)
condo_bw_mlr
```

```{r}
condo_bp_mlr <- ols_step_both_p(condo_mlr,
                                   p_val = 0.05,
                                   datails = FALSE)
condo_bp_mlr
```
model selection
```{r}
metric <- compare_performance(condo_mlr,
                              condo_fw_mlr$model,
                              condo_bw_mlr$model,
                              condo_bp_mlr$model
                )
```

gsub() is used to tidy the test value in name field
```{r}
metric$Name <- gsub(".*\\\\([a-zA-Z0-9_]+)\\\\,\\\\model\\\\.*","\\1",metric$Name)
```

comparing the performance
```{r}
plot(metric)
```

Visualising
```{r}
ggcoefstats(condo_bp_mlr$model,
            sort = "ascending")
```
checking formulticollinarity
```{r}
check_collinearity(condo_bp_mlr$model)
```
```{r}
plot(check_collinearity(condo_bp_mlr$model))+
  theme(axis.text = element_text(angle = 45,hjust = 1))
```
Linear the assumption test
```{R}
out <- plot(check_model(condo_bp_mlr$model,
                        panel = FALSE))
out[[2]]
```

Normality assumption test
```{r}
plot(check_normality(condo_bp_mlr$model))
```
checking the outline
```{r}
outliers <- check_outliers(condo_bp_mlr$model,
                           method = "cook")
outliers
```
```{r}
plot(check_outliers(condo_bp_mlr$model,
                           method = "cook"))
```

```{r}
mlr_output <- as.data.frame(condo_fw_mlr$model$residuals)%>%
  rename(`FW_MLR_RES`=`condo_fw_mlr$model$residuals`)
```

