---
title: "Take-home_Ex03"
author: "Yuheng Liang"
format: html
editor: visual
date: "Oct 28, 2024"
date-modified: "Oct 28, 2024"
execute: 
  eval: true
  echo: true
  freeze: true
---

# Take home exe 03

## Data

-   Philippines data from [Armed Conflict Location & Event Data Project (ACLED)](https://acleddata.com/).

-   [Philippines boundary data](https://data.humdata.org/dataset/cod-ab-phl) from HDX

## Importing data and package

### package

-   spNetwork, which provides functions to perform Spatial Point Patterns Analysis such as kernel density estimation (KDE) and K-function on network. It also can be used to build spatial matrices (‘listw’ objects like in ‘spdep’ package) to conduct any kind of traditional spatial analysis with spatial weights based on reticular distances.
-   sf package provides functions to manage, processing, and manipulate Simple Features, a formal geospatial data standard that specifies a storage and access model of spatial geometries such as points, lines, and polygons.
-   tmap which provides functions for plotting cartographic quality static point patterns maps or interactive maps by using leaflet API.
-   spatstat, which has a wide range of useful functions for point pattern analysis. In this hands-on exercise, it will be used to perform 1st- and 2nd-order spatial point patterns analysis and derive kernel density estimation (KDE) layer.
-   raster which reads, writes, manipulates, analyses and model of gridded spatial data (i.e. raster). In this hands-on exercise, it will be used to convert image output generate by spatstat into raster format.
-   maptools which provides a set of tools for manipulating geographic data. In this hands-on exercise, we mainly use it to convert Spatial objects into ppp format of spatstat.

```{r}
pacman::p_load(sf, raster, spatstat, tmap, tidyverse,sparr,spNetwork,dplyr,animation,stringr)
```

### data

```{r}
drug_case <- read_csv("data/2016-01-01-2024-06-30-Philippines.csv")%>%
    st_as_sf(coords = c("longitude", "latitude"), crs = 4326) %>%
    st_transform(crs = 32651) %>%
  mutate(event_date = dmy(event_date)) %>%
  mutate(event_month = year*100 + month(event_date)) %>%
  mutate(event_quarter = year*10 + quarter(event_date))%>%
  mutate(quarter = as.numeric(str_sub(event_quarter, 5, 5)))
```

import the boundary data

```{r}
ph_sf = st_read(dsn = "data/phl_adm_psa_namria_20231106_shp",layer = "phl_admbnda_adm2_psa_namria_20231106")
```

chack the crs

```{r}
st_crs(ph_sf)
```

transform the crs

```{r}
ph_sf = st_transform(ph_sf,crs = 32651)
st_crs(ph_sf)
```

```{r}
ph_sf_owin <-as.owin(ph_sf)
```

show the point

```{r}
tmap_mode('view')
tm_shape(drug_case)+tm_dots()
```

```{r}
tmap_mode('plot')
```

```{r}
hist(drug_case$year)
hist(drug_case$event_month)
```

```{r}
drug_case$fatalities <- as.numeric(drug_case$fatalities)

fatalities_by_year <- drug_case %>%
  group_by(year) %>%
  summarise(
    total_fatalities = sum(fatalities, na.rm = TRUE),
    average_fatalities = mean(fatalities, na.rm = TRUE),
    .groups = "drop"
  )

print(fatalities_by_year)
```

```{r}
cases_by_month <- drug_case %>%
  group_by(event_month) %>%
  summarise(case_count = n()) %>%
  arrange(event_month)

# 打印结果
print(cases_by_month)
```

```{r}

cases_by_quarter <- drug_case %>%
  group_by(event_quarter) %>%
  summarise(case_count = n()) %>%
  arrange(event_quarter)

print(cases_by_quarter)
```

```{r}
ggplot(cases_by_month, aes(x = event_month, y = case_count)) +
  geom_bar(stat = "identity", fill = "skyblue") +
  labs(title = "case_month", x = "month", y = "num") +
  theme_minimal()

ggplot(cases_by_quarter, aes(x = event_quarter, y = case_count)) +
  geom_bar(stat = "identity", fill = "lightgreen") +
  labs(title = "case by q", x = "q", y = "num") +
  theme_minimal()
```

Excluding the third and fourth quarters of 2016 and the first two or three quarters of 2017 and the first two or two quarters of 2021 the distribution of the other years is fairly even not described as seasonal \## Spatial Point Analysis by year Screening of drug use data for each year

## Spatio-Temporal Point Pattern Analysis by year
```{r}
drug_case_2016 <- drug_case %>%
  filter(year== 2016)
drug_case_2017 <- drug_case %>%
  filter(year== 2017)
drug_case_2018 <- drug_case %>%
  filter(year== 2018)
drug_case_2019 <- drug_case %>%
  filter(year== 2019)
drug_case_2020 <- drug_case %>%
  filter(year== 2020)
drug_case_2021 <- drug_case %>%
  filter(year== 2021)
drug_case_2022 <- drug_case %>%
  filter(year== 2022)
drug_case_2023 <- drug_case %>%
  filter(year== 2023)
drug_case_2024 <- drug_case %>%
  filter(year== 2024)
```

transform them into ppp objects.

```{r}
drug_case_2016_ppp <- as.ppp(drug_case_2016)
drug_case_2017_ppp <- as.ppp(drug_case_2017)
drug_case_2018_ppp <- as.ppp(drug_case_2018)
drug_case_2019_ppp <- as.ppp(drug_case_2019)
drug_case_2020_ppp <- as.ppp(drug_case_2020)
drug_case_2021_ppp <- as.ppp(drug_case_2021)
drug_case_2022_ppp <- as.ppp(drug_case_2022)
drug_case_2023_ppp <- as.ppp(drug_case_2023)
drug_case_2024_ppp <- as.ppp(drug_case_2024)
```

```{r}
drug_case_2016_ppp <- drug_case_2016_ppp[ph_sf_owin]
drug_case_2017_ppp <- drug_case_2017_ppp[ph_sf_owin]
drug_case_2018_ppp <- drug_case_2018_ppp[ph_sf_owin]
drug_case_2019_ppp <- drug_case_2019_ppp[ph_sf_owin]
drug_case_2020_ppp <- drug_case_2020_ppp[ph_sf_owin]
drug_case_2021_ppp <- drug_case_2021_ppp[ph_sf_owin]
drug_case_2022_ppp <- drug_case_2022_ppp[ph_sf_owin]
drug_case_2023_ppp <- drug_case_2023_ppp[ph_sf_owin]
drug_case_2024_ppp <- drug_case_2024_ppp[ph_sf_owin]
```

re-scale them to km

```{r}
drug_case_2016_ppp.km <- rescale.ppp(drug_case_2016_ppp, 1000, "km")
drug_case_2017_ppp.km <- rescale.ppp(drug_case_2017_ppp, 1000, "km")
drug_case_2018_ppp.km <- rescale.ppp(drug_case_2018_ppp, 1000, "km")
drug_case_2019_ppp.km <- rescale.ppp(drug_case_2019_ppp, 1000, "km")
drug_case_2020_ppp.km <- rescale.ppp(drug_case_2020_ppp, 1000, "km")
drug_case_2021_ppp.km <- rescale.ppp(drug_case_2021_ppp, 1000, "km")
drug_case_2022_ppp.km <- rescale.ppp(drug_case_2022_ppp, 1000, "km")
drug_case_2023_ppp.km <- rescale.ppp(drug_case_2023_ppp, 1000, "km")
drug_case_2024_ppp.km <- rescale.ppp(drug_case_2024_ppp, 1000, "km")
```

drug_case_2016_ppp_owin.csr <- envelope(drug_case_2016_ppp.km, Gest, nsim = 99)
drug_case_2017_ppp_owin.csr <- envelope(drug_case_2017_ppp.km, Gest, nsim = 99)
drug_case_2018_ppp_owin.csr <- envelope(drug_case_2018_ppp.km, Gest, nsim = 99)
drug_case_2019_ppp_owin.csr <- envelope(drug_case_2019_ppp.km, Gest, nsim = 99)
drug_case_2020_ppp_owin.csr <- envelope(drug_case_2020_ppp.km, Gest, nsim = 99)
drug_case_2021_ppp_owin.csr <- envelope(drug_case_2021_ppp.km, Gest, nsim = 99)
drug_case_2022_ppp_owin.csr <- envelope(drug_case_2022_ppp.km, Gest, nsim = 99)
drug_case_2023_ppp_owin.csr <- envelope(drug_case_2023_ppp.km, Gest, nsim = 99)
drug_case_2024_ppp_owin.csr <- envelope(drug_case_2024_ppp.km, Gest, nsim = 99)

plot(drug_case_2016_ppp_owin.csr)
plot(drug_case_2017_ppp_owin.csr)
plot(drug_case_2018_ppp_owin.csr)
plot(drug_case_2019_ppp_owin.csr)
plot(drug_case_2020_ppp_owin.csr)
plot(drug_case_2021_ppp_owin.csr)
plot(drug_case_2022_ppp_owin.csr)
plot(drug_case_2023_ppp_owin.csr)
plot(drug_case_2024_ppp_owin.csr)

## Spatio-Temporal Point Pattern Analysis by quarter


```{r}
drug_case_2016_q1 <- drug_case %>%
  filter(year== 2016,event_quarter == 20161)
drug_case_2016_q2 <- drug_case %>%
  filter(year== 2016,event_quarter == 20162)
drug_case_2016_q3 <- drug_case %>%
  filter(year== 2016,event_quarter == 20163)
drug_case_2016_q4 <- drug_case %>%
  filter(year== 2016,event_quarter == 20164)
```

```{R}
drug_case_2017_q1 <- drug_case %>%
  filter(year== 2017,event_quarter == 20171)
drug_case_2017_q2 <- drug_case %>%
  filter(year== 2017,event_quarter == 20172)
drug_case_2017_q3 <- drug_case %>%
  filter(year== 2017,event_quarter == 20173)
drug_case_2017_q4 <- drug_case %>%
  filter(year== 2017,event_quarter == 20174)
```

```{R}
drug_case_2018_q1 <- drug_case %>%
  filter(year== 2018,event_quarter == 20181)
drug_case_2018_q2 <- drug_case %>%
  filter(year== 2018,event_quarter == 20182)
drug_case_2018_q3 <- drug_case %>%
  filter(year== 2018,event_quarter == 20183)
drug_case_2018_q4 <- drug_case %>%
  filter(year== 2018,event_quarter == 20184)
```

```{r}
drug_case_2019_q1 <- drug_case %>%
  filter(year== 2019,event_quarter == 20191)
drug_case_2019_q2 <- drug_case %>%
  filter(year== 2019,event_quarter == 20192)
drug_case_2019_q3 <- drug_case %>%
  filter(year== 2019,event_quarter == 20193)
drug_case_2019_q4 <- drug_case %>%
  filter(year== 2019,event_quarter == 20194)
```

```{r}
drug_case_2020_q1 <- drug_case %>%
  filter(year== 2020,event_quarter == 20201)
drug_case_2020_q2 <- drug_case %>%
  filter(year== 2020,event_quarter == 20202)
drug_case_2020_q3 <- drug_case %>%
  filter(year== 2020,event_quarter == 20203)
drug_case_2020_q4 <- drug_case %>%
  filter(year== 2020,event_quarter == 20204)
```

```{R}
drug_case_2021_q1 <- drug_case %>%
  filter(year== 2021,event_quarter == 20211)
drug_case_2021_q2 <- drug_case %>%
  filter(year== 2021,event_quarter == 20212)
drug_case_2021_q3 <- drug_case %>%
  filter(year== 2021,event_quarter == 20213)
drug_case_2021_q4 <- drug_case %>%
  filter(year== 2021,event_quarter == 20214)
```

```{R}
drug_case_2022_q1 <- drug_case%>%
  filter(year== 2022,event_quarter == 20221)
drug_case_2022_q2 <- drug_case%>%
  filter(year== 2022,event_quarter == 20222)
drug_case_2022_q3 <- drug_case%>%
  filter(year== 2022,event_quarter == 20223)
drug_case_2022_q4 <- drug_case%>%
  filter(year== 2022,event_quarter == 20224)
```

```{r}
drug_case_2023_q1 <- drug_case%>%
  filter(year== 2023,event_quarter == 20231)
drug_case_2023_q2 <- drug_case%>%
  filter(year== 2023,event_quarter == 20232)
drug_case_2023_q3 <- drug_case%>%
  filter(year== 2023,event_quarter == 20233)
drug_case_2023_q4 <- drug_case%>%
  filter(year== 2023,event_quarter == 20234)
```

```{r}
drug_case_2024_q1 <- drug_case%>%
  filter(year== 2024,event_quarter == 20241)
drug_case_2024_q2 <- drug_case%>%
  filter(year== 2024,event_quarter == 20242)
```

```{r}
drug_case_2016_ppp1 <- as.ppp(drug_case_2016_q1)
drug_case_2016_ppp2 <- as.ppp(drug_case_2016_q2)
drug_case_2016_ppp3 <- as.ppp(drug_case_2016_q3)
drug_case_2016_ppp4 <- as.ppp(drug_case_2016_q4)
```

```{r}
drug_case_2017_ppp1 <- as.ppp(drug_case_2017_q1)
drug_case_2017_ppp2 <- as.ppp(drug_case_2017_q2)
drug_case_2017_ppp3 <- as.ppp(drug_case_2017_q3)
drug_case_2017_ppp4 <- as.ppp(drug_case_2017_q4)
```

```{R}
drug_case_2018_ppp1 <- as.ppp(drug_case_2018_q1)
drug_case_2018_ppp2 <- as.ppp(drug_case_2018_q2)
drug_case_2018_ppp3 <- as.ppp(drug_case_2018_q3)
drug_case_2018_ppp4 <- as.ppp(drug_case_2018_q4)
```

```{r}
drug_case_2019_ppp1 <- as.ppp(drug_case_2019_q1)
drug_case_2019_ppp2 <- as.ppp(drug_case_2019_q2)
drug_case_2019_ppp3 <- as.ppp(drug_case_2019_q3)
drug_case_2019_ppp4 <- as.ppp(drug_case_2019_q4)
```

```{r}
drug_case_2020_ppp1 <- as.ppp(drug_case_2020_q1)
drug_case_2020_ppp2 <- as.ppp(drug_case_2020_q2)
drug_case_2020_ppp3 <- as.ppp(drug_case_2020_q3)
drug_case_2020_ppp4 <- as.ppp(drug_case_2020_q4)
```

```{R}
drug_case_2021_ppp1 <- as.ppp(drug_case_2021_q1)
drug_case_2021_ppp2 <- as.ppp(drug_case_2021_q2)
drug_case_2021_ppp3 <- as.ppp(drug_case_2021_q3)
drug_case_2021_ppp4 <- as.ppp(drug_case_2021_q4)
```

```{r}
drug_case_2022_ppp1 <- as.ppp(drug_case_2022_q1)
drug_case_2022_ppp2 <- as.ppp(drug_case_2022_q2)
drug_case_2022_ppp3 <- as.ppp(drug_case_2022_q3)
drug_case_2022_ppp4 <- as.ppp(drug_case_2022_q4)
```

```{r}
drug_case_2023_ppp1 <- as.ppp(drug_case_2023_q1)
drug_case_2023_ppp2 <- as.ppp(drug_case_2023_q2)
drug_case_2023_ppp3 <- as.ppp(drug_case_2023_q3)
drug_case_2023_ppp4 <- as.ppp(drug_case_2023_q4)
```

```{R}
drug_case_2024_ppp1 <- as.ppp(drug_case_2024_q1)
drug_case_2024_ppp2 <- as.ppp(drug_case_2024_q2)
```

```{r}
drug_case_2016_ppp1_owin <- drug_case_2016_ppp1[ph_sf_owin]
drug_case_2016_ppp2_owin <- drug_case_2016_ppp2[ph_sf_owin]
drug_case_2016_ppp3_owin <- drug_case_2016_ppp3[ph_sf_owin]
drug_case_2016_ppp4_owin <- drug_case_2016_ppp4[ph_sf_owin]
```

```{r}
drug_case_2017_ppp1_owin <- drug_case_2017_ppp1[ph_sf_owin]
drug_case_2017_ppp2_owin <- drug_case_2017_ppp2[ph_sf_owin]
drug_case_2017_ppp3_owin <- drug_case_2017_ppp3[ph_sf_owin]
drug_case_2017_ppp4_owin <- drug_case_2017_ppp4[ph_sf_owin]
```

```{R}
drug_case_2018_ppp1_owin <- drug_case_2018_ppp1[ph_sf_owin]
drug_case_2018_ppp2_owin <- drug_case_2018_ppp2[ph_sf_owin]
drug_case_2018_ppp3_owin <- drug_case_2018_ppp3[ph_sf_owin]
drug_case_2018_ppp4_owin <- drug_case_2018_ppp4[ph_sf_owin]
```

```{R}
drug_case_2019_ppp1_owin <- drug_case_2019_ppp1[ph_sf_owin]
drug_case_2019_ppp2_owin <- drug_case_2019_ppp2[ph_sf_owin]
drug_case_2019_ppp3_owin <- drug_case_2019_ppp3[ph_sf_owin]
drug_case_2019_ppp4_owin <- drug_case_2019_ppp4[ph_sf_owin]
```

```{r}
drug_case_2020_ppp1_owin <- drug_case_2020_ppp1[ph_sf_owin]
drug_case_2020_ppp2_owin <- drug_case_2020_ppp2[ph_sf_owin]
drug_case_2020_ppp3_owin <- drug_case_2020_ppp3[ph_sf_owin]
drug_case_2020_ppp4_owin <- drug_case_2020_ppp4[ph_sf_owin]
```

```{R}
drug_case_2021_ppp1_owin <- drug_case_2021_ppp1[ph_sf_owin]
drug_case_2021_ppp2_owin <- drug_case_2021_ppp2[ph_sf_owin]
drug_case_2021_ppp3_owin <- drug_case_2021_ppp3[ph_sf_owin]
drug_case_2021_ppp4_owin <- drug_case_2021_ppp4[ph_sf_owin]
```

```{R}
drug_case_2022_ppp1_owin <- drug_case_2022_ppp1[ph_sf_owin]
drug_case_2022_ppp2_owin <- drug_case_2022_ppp2[ph_sf_owin]
drug_case_2022_ppp3_owin <- drug_case_2022_ppp3[ph_sf_owin]
drug_case_2022_ppp4_owin <- drug_case_2022_ppp4[ph_sf_owin]
```

```{R}
drug_case_2023_ppp1_owin <- drug_case_2023_ppp1[ph_sf_owin]
drug_case_2023_ppp2_owin <- drug_case_2023_ppp2[ph_sf_owin]
drug_case_2023_ppp3_owin <- drug_case_2023_ppp3[ph_sf_owin]
drug_case_2023_ppp4_owin <- drug_case_2023_ppp4[ph_sf_owin]
```

```{r}
drug_case_2024_ppp1_owin <- drug_case_2024_ppp1[ph_sf_owin]
drug_case_2024_ppp2_owin <- drug_case_2024_ppp2[ph_sf_owin]
```

```{r}
drug_case_2016_ppp1.km <- rescale.ppp(drug_case_2016_ppp1_owin, 1000, "km")
drug_case_2016_ppp2.km <- rescale.ppp(drug_case_2016_ppp2_owin, 1000, "km")
drug_case_2016_ppp3.km <- rescale.ppp(drug_case_2016_ppp3_owin, 1000, "km")
drug_case_2016_ppp4.km <- rescale.ppp(drug_case_2016_ppp4_owin, 1000, "km")
```

```{r}
drug_case_2017_ppp1.km <- rescale.ppp(drug_case_2017_ppp1_owin, 1000, "km")
drug_case_2017_ppp2.km <- rescale.ppp(drug_case_2017_ppp2_owin, 1000, "km")
drug_case_2017_ppp3.km <- rescale.ppp(drug_case_2017_ppp3_owin, 1000, "km")
drug_case_2017_ppp4.km <- rescale.ppp(drug_case_2017_ppp4_owin, 1000, "km")
```

```{r}
drug_case_2018_ppp1.km <- rescale.ppp(drug_case_2018_ppp1_owin, 1000, "km")
drug_case_2018_ppp2.km <- rescale.ppp(drug_case_2018_ppp2_owin, 1000, "km")
drug_case_2018_ppp3.km <- rescale.ppp(drug_case_2018_ppp3_owin, 1000, "km")
drug_case_2018_ppp4.km <- rescale.ppp(drug_case_2018_ppp4_owin, 1000, "km")
```

```{R}
drug_case_2019_ppp1.km <- rescale.ppp(drug_case_2019_ppp1_owin, 1000, "km")
drug_case_2019_ppp2.km <- rescale.ppp(drug_case_2019_ppp2_owin, 1000, "km")
drug_case_2019_ppp3.km <- rescale.ppp(drug_case_2019_ppp3_owin, 1000, "km")
drug_case_2019_ppp4.km <- rescale.ppp(drug_case_2019_ppp4_owin, 1000, "km")
```

```{r}
drug_case_2020_ppp1.km <- rescale.ppp(drug_case_2020_ppp1_owin, 1000, "km")
drug_case_2020_ppp2.km <- rescale.ppp(drug_case_2020_ppp2_owin, 1000, "km")
drug_case_2020_ppp3.km <- rescale.ppp(drug_case_2020_ppp3_owin, 1000, "km")
drug_case_2020_ppp4.km <- rescale.ppp(drug_case_2020_ppp4_owin, 1000, "km")
```

```{r}
drug_case_2021_ppp1.km <- rescale.ppp(drug_case_2021_ppp1_owin, 1000, "km")
drug_case_2021_ppp2.km <- rescale.ppp(drug_case_2021_ppp2_owin, 1000, "km")
drug_case_2021_ppp3.km <- rescale.ppp(drug_case_2021_ppp3_owin, 1000, "km")
drug_case_2021_ppp4.km <- rescale.ppp(drug_case_2021_ppp4_owin, 1000, "km")
```

```{r}
drug_case_2022_ppp1.km <- rescale.ppp(drug_case_2022_ppp1_owin, 1000, "km")
drug_case_2022_ppp2.km <- rescale.ppp(drug_case_2022_ppp2_owin, 1000, "km")
drug_case_2022_ppp3.km <- rescale.ppp(drug_case_2022_ppp3_owin, 1000, "km")
drug_case_2022_ppp4.km <- rescale.ppp(drug_case_2022_ppp4_owin, 1000, "km")
```

```{r}
drug_case_2023_ppp1.km <- rescale.ppp(drug_case_2023_ppp1_owin, 1000, "km")
drug_case_2023_ppp2.km <- rescale.ppp(drug_case_2023_ppp2_owin, 1000, "km")
drug_case_2023_ppp3.km <- rescale.ppp(drug_case_2023_ppp3_owin, 1000, "km")
drug_case_2023_ppp4.km <- rescale.ppp(drug_case_2023_ppp4_owin, 1000, "km")
```

```{r}
drug_case_2024_ppp1.km <- rescale.ppp(drug_case_2024_ppp1_owin, 1000, "km")
drug_case_2024_ppp2.km <- rescale.ppp(drug_case_2024_ppp2_owin, 1000, "km")
```


drug_case_2016_ppp1.km.csr <- envelope(drug_case_2016_ppp1.km, Gest, nsim = 99)
drug_case_2016_ppp2.km.csr <- envelope(drug_case_2016_ppp1.km, Gest, nsim = 99)
drug_case_2016_ppp3.km.csr <- envelope(drug_case_2016_ppp1.km, Gest, nsim = 99)
drug_case_2016_ppp4.km.csr <- envelope(drug_case_2016_ppp1.km, Gest, nsim = 99)
drug_case_2017_ppp1.km.csr <- envelope(drug_case_2017_ppp1.km, Gest, nsim = 99)
drug_case_2018_ppp2.km.csr <- envelope(drug_case_2018_ppp1.km, Gest, nsim = 99)
drug_case_2019_ppp3.km.csr <- envelope(drug_case_2019_ppp1.km, Gest, nsim = 99)
drug_case_2020_ppp4.km.csr <- envelope(drug_case_2020_ppp1.km, Gest, nsim = 99)
drug_case_2021_ppp1.km.csr <- envelope(drug_case_2021_ppp1.km, Gest, nsim = 99)
drug_case_2022_ppp1.km.csr <- envelope(drug_case_2022_ppp1.km, Gest, nsim = 99)
drug_case_2023_ppp1.km.csr <- envelope(drug_case_2023_ppp1.km, Gest, nsim = 99)
drug_case_2024_ppp1.km.csr <- envelope(drug_case_2024_ppp1.km, Gest, nsim = 99)

plot(drug_case_2016_ppp1.km.csr)


## Spatio-Temporal Point Pattern Analysis by month
```{r}
drug_case_2016_1 <- drug_case %>%
  filter(year== 2016, event_month == 201601)
```

transform them into ppp objects.
```{r}
drug_case_2016_1_ppp <- as.ppp(drug_case_2016_1)
```

combine
```{r}
drug_case_2016_1_ppp <- drug_case_2016_1_ppp[ph_sf_owin]
```

```{r}
drug_case_2016_1_ppp.km <- rescale.ppp(drug_case_2016_1_ppp, 1000, "km")
```

drug_case_2016_ppp_owin.csr <- envelope(drug_case_2016_ppp.km, Gest, nsim = 99)
plot(drug_case_2016_ppp_owin.csr)
Filtering data


## Spatio-Temporal Point Pattern Analysis by types

Filtering data

```{r}
Violence_against_civilians <- drug_case%>%
  filter(event_type == "Violence against civilians")
Strategic_developments <- drug_case%>%
  filter(event_type == "Strategic developments")
Battles <- drug_case%>%
  filter(event_type == "Battles")
```


### Violence against civilians by year

Filtering data

```{r}
Violence_16 = Violence_against_civilians %>% 
  filter(year == 2016) %>% 
  dplyr::select(quarter)
Violence_17 = Violence_against_civilians %>% 
  filter(year == 2017) %>% 
  dplyr::select(quarter)
Violence_18 = Violence_against_civilians %>% 
  filter(year == 2018) %>% 
  dplyr::select(quarter)
Violence_19 = Violence_against_civilians %>% 
  filter(year == 2019) %>% 
  dplyr::select(quarter)
Violence_20 = Violence_against_civilians %>% 
  filter(year == 2020) %>% 
  dplyr::select(quarter)
Violence_21 = Violence_against_civilians %>% 
  filter(year == 2021) %>% 
  dplyr::select(quarter)
Violence_22 = Violence_against_civilians %>% 
  filter(year == 2022) %>% 
  dplyr::select(quarter)
Violence_23 = Violence_against_civilians %>% 
  filter(year == 2023) %>% 
  dplyr::select(quarter)
Violence_24 = Violence_against_civilians %>% 
  filter(year == 2024) %>% 
  dplyr::select(quarter)
```

```{r}
Violence_16_ppp<- as.ppp(Violence_16)
Violence_17_ppp<- as.ppp(Violence_17)
Violence_18_ppp<- as.ppp(Violence_18)
Violence_19_ppp<- as.ppp(Violence_19)
Violence_20_ppp<- as.ppp(Violence_20)
Violence_21_ppp<- as.ppp(Violence_21)
Violence_22_ppp<- as.ppp(Violence_22)
Violence_23_ppp<- as.ppp(Violence_23)
Violence_24_ppp<- as.ppp(Violence_24)
```

```{r}
Violence_16_ppp = Violence_16_ppp[ph_sf_owin]
Violence_17_ppp = Violence_17_ppp[ph_sf_owin]
Violence_18_ppp = Violence_18_ppp[ph_sf_owin]
Violence_19_ppp = Violence_19_ppp[ph_sf_owin]
Violence_20_ppp = Violence_20_ppp[ph_sf_owin]
Violence_21_ppp = Violence_21_ppp[ph_sf_owin]
Violence_22_ppp = Violence_22_ppp[ph_sf_owin]
Violence_23_ppp = Violence_23_ppp[ph_sf_owin]
Violence_24_ppp = Violence_24_ppp[ph_sf_owin]
```

Violence_16_ppp.csr = envelope(Violence_16_ppp, Gest, nsim = 99)

### Battles

Filtering data

```{r}
Battles_16 = Battles %>% 
  filter(year == 2016) %>% 
  dplyr::select(quarter)
Battles_17 = Battles %>% 
  filter(year == 2017) %>% 
  dplyr::select(quarter)
Battles_18 = Battles %>% 
  filter(year == 2018) %>% 
  dplyr::select(quarter)
Battles_19 = Battles %>% 
  filter(year == 2019) %>% 
  dplyr::select(quarter)
Battles_20 = Battles %>% 
  filter(year == 2020) %>% 
  dplyr::select(quarter)
Battles_21 = Battles %>% 
  filter(year == 2021) %>% 
  dplyr::select(quarter)
Battles_22 = Battles %>% 
  filter(year == 2022) %>% 
  dplyr::select(quarter)
Battles_23 = Battles %>% 
  filter(year == 2023) %>% 
  dplyr::select(quarter)
Battles_24 = Battles %>% 
  filter(year == 2024) %>% 
  dplyr::select(quarter)
```

```{r}
Battles_16_ppp<- as.ppp(Battles_16)
Battles_17_ppp<- as.ppp(Battles_17)
Battles_18_ppp<- as.ppp(Battles_18)
Battles_19_ppp<- as.ppp(Battles_19)
Battles_20_ppp<- as.ppp(Battles_20)
Battles_21_ppp<- as.ppp(Battles_21)
Battles_22_ppp<- as.ppp(Battles_22)
Battles_23_ppp<- as.ppp(Battles_23)
Battles_24_ppp<- as.ppp(Battles_24)
```

```{r}
Battles_16_ppp = Battles_16_ppp[ph_sf_owin]
Battles_17_ppp = Battles_17_ppp[ph_sf_owin]
Battles_18_ppp = Battles_18_ppp[ph_sf_owin]
Battles_19_ppp = Battles_19_ppp[ph_sf_owin]
Battles_20_ppp = Battles_20_ppp[ph_sf_owin]
Battles_21_ppp = Battles_21_ppp[ph_sf_owin]
Battles_22_ppp = Battles_22_ppp[ph_sf_owin]
Battles_23_ppp = Battles_23_ppp[ph_sf_owin]
Battles_24_ppp = Battles_24_ppp[ph_sf_owin]
```

Battles_16_ppp.csr = envelope(Battles_16_ppp, Gest, nsim = 99)

### Strategic developments

```{r}
Strategic_developments_16 = Strategic_developments %>% 
  filter(year == 2016) %>% 
  dplyr::select(quarter)
Strategic_developments_17 = Strategic_developments %>% 
  filter(year == 2017) %>% 
  dplyr::select(quarter)
Strategic_developments_18 = Strategic_developments %>% 
  filter(year == 2018) %>% 
  dplyr::select(quarter)
Strategic_developments_19 = Strategic_developments %>% 
  filter(year == 2019) %>% 
  dplyr::select(quarter)
Strategic_developments_20 = Strategic_developments %>% 
  filter(year == 2020) %>% 
  dplyr::select(quarter)
Strategic_developments_21 = Strategic_developments %>% 
  filter(year == 2021) %>% 
  dplyr::select(quarter)
Strategic_developments_22 = Strategic_developments %>% 
  filter(year == 2022) %>% 
  dplyr::select(quarter)
Strategic_developments_23 = Strategic_developments %>% 
  filter(year == 2023) %>% 
  dplyr::select(quarter)
Strategic_developments_24 = Strategic_developments %>% 
  filter(year == 2024) %>% 
  dplyr::select(quarter)
```

```{r}
Strategic_developments_16_ppp <- as.ppp(Strategic_developments_16)
Strategic_developments_17_ppp <- as.ppp(Strategic_developments_17)
Strategic_developments_18_ppp <- as.ppp(Strategic_developments_18)
Strategic_developments_19_ppp <- as.ppp(Strategic_developments_19)
Strategic_developments_20_ppp <- as.ppp(Strategic_developments_20)
Strategic_developments_21_ppp <- as.ppp(Strategic_developments_21)
Strategic_developments_22_ppp <- as.ppp(Strategic_developments_22)
Strategic_developments_23_ppp <- as.ppp(Strategic_developments_23)
Strategic_developments_24_ppp <- as.ppp(Strategic_developments_24)
```

```{r}
Strategic_developments_16_ppp = Strategic_developments_16_ppp[ph_sf_owin]
Strategic_developments_17_ppp = Strategic_developments_17_ppp[ph_sf_owin]
Strategic_developments_18_ppp = Strategic_developments_18_ppp[ph_sf_owin]
Strategic_developments_19_ppp = Strategic_developments_19_ppp[ph_sf_owin]
Strategic_developments_20_ppp = Strategic_developments_20_ppp[ph_sf_owin]
Strategic_developments_21_ppp = Strategic_developments_21_ppp[ph_sf_owin]
Strategic_developments_22_ppp = Strategic_developments_22_ppp[ph_sf_owin]
Strategic_developments_23_ppp = Strategic_developments_23_ppp[ph_sf_owin]
Strategic_developments_24_ppp = Strategic_developments_24_ppp[ph_sf_owin]
```

Strategic_developments_16_ppp.csr = envelope(Strategic_developments_16_ppp, Gest, nsim = 99)
