---
title: "In-class_Ex04"
author: "Yuheng Liang"
format: html
editor: visual
date: "September 9, 2024"
date-modified: "September 9,2024"
execute: 
  eval: true
  echo: true
  freeze: true
---

# In-class Ex01

## Data

### Installing package

-   sf

-   raster

-   spatstat

-   sparr

-   tmap

-   tidyverse

```{R}
pacman::p_load(sf, raster, spatstat,sparr,tmap, tidyverse)
```
### Importing Data
#### Importing study area
```{R}
kbb <- st_read(dsn="data/rawdata", layer="Kepulauan_Bangka_Belitung") %>%
st_transform(crs=32748)
```

```{r}
st_crs(kbb)
```

the revised chock chunk
```{R}
kbb_sf <- st_read(dsn="data/rawdata", layer="Kepulauan_Bangka_Belitung") %>%
  st_union()%>%
  st_zm(drop= TRUE, waht ="ZM")%>%
  st_transform(crs=32748)
```
#### Converting OWIN
Next, as.owin is used to convert kbb into an owin object
```{R}
kbb_owin <- as.owin(kbb_sf)
kbb_owin
```
check data
```{r}
class(kbb_owin)
```

### Importing forest fire data
```{R}
fire_sf <- read_csv("data/rawdata/forestfires.csv")%>%
  st_as_sf(coords =c("longitude","latitude"),crs=4326)%>%
  st_transform(crs=32748)
```

```{R}
class(fire_sf)
```
Because ppp only accept numerical or character as mark.
```{R}
fire_sf <- fire_sf%>%
  mutate(DayofYear = yday(acq_date))%>%
  mutate(Month_num = month(acq_date))%>%
  mutate(Month_fac = month(acq_date, label=TRUE, abbr = FALSE))
```

## Visualising the Fire point
### Overall pllot
```{R}
tm_shape(kbb_sf)+tm_polygons()+tm_shape(fire_sf)+tm_dots()
```

### visualsing geographiv distribution of forest fire by month
```{R}
tm_shape(kbb_sf)+tm_polygons()+tm_shape(fire_sf)+tm_dots(size =0.1)+ tm_facets(by ="Month_fac", free.coords = FALSE, drop.units = TRUE )
```
## Computing Stkde by month
### Extracting forest fires by month
```{R}
fire_month <-fire_sf %>%
  select(Month_num)
```

### Creating PPP
```{R}
fire_month_ppp <- as.ppp(fire_month)
fire_month_ppp
```
check the output is in the correct object class
```{r}
summary(fire_month_ppp)
```
```{r}
any(duplicated(fire_month_ppp))
```

### Including Owin object
combine they into one
```{R}
fire_month_owin <-fire_month_ppp[kbb_owin]
summary(fire_month_owin)
```
### Computing spatio-temporal KDE
Compute the STKDE
```{R}
st_kde <- spattemp.density(fire_month_owin)
summary(st_kde)
```

### Ploting the Spatio-temporal KDE object
```{R}
tims <- c(7,8,9,10,11,12)
par(mfcol= c(2,3))
for(i in tims){
  plot(st_kde,i,  
  override.par= FALSE,
  fix.range = TRUE,
  main=paste("KDE AT MONTH", i))
}
```




